# Lab 3: `coronavirus` visualization, data wrangling, and dates

The package is available on GitHub [here](https://github.com/RamiKrispin/coronavirus) and is updated daily.

> I use the `coronavirus` package and use the `coronavirus::update_data()` function to keep the data current.


## Let'slook like Applied Analytics Superstars and make some neat visuals.

 

```{r setup, message=FALSE, warning=FALSE}
library(coronavirus)
head(coronavirus)
library(dplyr)
library(ggplot2)
```

I'd recommend you always start by trying to understand a bit about the data.

```{r}
head(coronavirus)

```

For exmple, what does this summary let us know?

```{r}
summary(coronavirus$cases)
```

1. Can you create a visual showing the cases over time for China, Russia, Spain, US, and Venezuela?
Also, why might `filter(cases >= 0)` be worth using? 

```{r echo=FALSE, message=FALSE, warning=FALSE}
cases.over.time = coronavirus %>%
  filter(country == "Russia"| 
           country == "Venezuela"| country == "US"| country == "Spain") %>%
  filter(cases >= 0) %>%
  group_by(country,date) %>%
  summarize(total.cases = sum(cases))

ggplot(data = cases.over.time, mapping = aes(x=date, y=total.cases, color = country)) +
  geom_line() +
  labs(title = "Cases Over Time in 4 Countries",
       subtitle = "Russia, Spain, US, and Venezuela",
       y = "Cases",
       x = "Date")
```

2. Can you show deaths over time for Russia, Spain, US, and Venezuela?  And can you play with your geoms and make something neat?

```{r message=FALSE, warning=FALSE, echo=FALSE}
deaths.over.time = coronavirus %>%
  filter(country == "China"| country == "Russia"| 
           country == "Venezuela"| country == "US"| country == "Spain") %>%
  filter(type == "death") %>%
  filter(cases >= 0) %>%
  group_by(country,date) %>%
  summarize(total.deaths = sum(cases))

ggplot(data = deaths.over.time, mapping = aes(x=date, y=total.deaths, color = country)) +
  geom_smooth() +
  geom_line() +
  labs(title = "Deaths Over Time in 4 Countries",
       x = "Date",
       y = "COVID deaths")
```

3. Now lets do a plot of COVID rate (# cases / population).  Something like this. 

```{r message=FALSE, warning=FALSE, echo=FALSE}
rate.over.time = coronavirus %>%
  filter(country == "China"| country == "Russia"| 
           country == "Venezuela"| country == "US"| country == "Spain") %>%
  filter(cases >= 0) %>%
  group_by(country,date,population) %>%
  summarize(total.cases = sum(cases)) %>%
  mutate(covid.rate = total.cases/population)

ggplot(data = rate.over.time, mapping = aes(x=date, y=covid.rate, color = country)) +
  geom_smooth() +
  labs(title = "Covid Rate in the Population of 5 Different Countries",
       x = "Date",
       y = "Covid Rate (Cases/Population)")

```

4. What is not useful about the previous illustration?  

5. Make a chart with cumulative cases.  Something like this:

```{r essage=FALSE, warning=FALSE, echo=FALSE}
cases <- coronavirus %>% 
  filter(country=="China"|country=="Russia"|country=="Spain"|country=="Venezuela"|country=="US") %>%
  filter(type=="confirmed") %>% filter(cases>!0) %>% 
  select(date,country, cases,type) %>% group_by(date,country) %>% 
  mutate("total cases"= sum(cases)) %>% 
  distinct(date,country, .keep_all=TRUE) %>% 
  select(date,country,cases,"total cases") %>% 
  group_by(country) %>% 
  mutate("cumulative cases" = cumsum(cases)) %>% 
  select(date,country, "total cases", "cumulative cases") 

ggplot(data = cases, mapping = aes(date,`cumulative cases`, color=country))+
  geom_line(size= 1)+
  labs(x="Date", y= "Cases", 
  title = "Covid Cases", 
  subtitle = "Cumulative covid cases by country")
```

6.  Now let's **really** have some fun.  Let's illustrate death rates relative to confirmed cases.  Why is this more challenging than anything we've done so far in this lab?  We're going to have to make this data **tidy**.

7. One way to play this game.

knitr::opts_chunk$set(echo = FALSE)

```{r message=FALSE, warning=FALSE, echo=FALSE}
df1 <- coronavirus %>% 
  filter(country=="Russia"|country=="US"|country=="Spain"|country=="Venezuela", type == "death") %>%
    filter(cases >= 0) %>%
  rename(deaths = cases) %>% 
  select(date, country, deaths)
```

Let's make a little table of just deaths.

```{r message=FALSE, warning=FALSE, echo=FALSE}
head(df1)
df1 %>% count(country)
```

Let's make a little table of just confirmed cases.

```{r message=FALSE, warning=FALSE, echo=FALSE}
df2 <- coronavirus %>% 
  filter(country=="Russia"|country=="US"|country=="Spain"|country=="Venezuela", type == "confirmed") %>% 
  rename(confirmed = cases) %>% 
  select(date, country, confirmed)

df2 %>% count(country)
head(df2)

```

Let's join these together. I use `left_join`.  


```{r message=FALSE, warning=FALSE, echo=FALSE}
df3 <- full_join(df1, df2, keep = FALSE)
head(df3)
df3 %>% count(country)
df3 <- df3 %>% 
  #  group_by(country,date) %>% 
  #  summarise(tot_cases = sum(cases)) %>% 
  arrange(country, date) %>% 
  mutate(cumulative_cases = cumsum(confirmed)) %>% 
  mutate(cumulative_deaths = cumsum(deaths))  %>% 
  mutate(rate = if_else(confirmed == 0, 0, (deaths/confirmed)))


ggplot(data = df3,
       mapping = aes(x = date,
                     y = cumulative_cases,
                     color = country)) +
  geom_line()


summary(df3$cumulative_deaths)
goofy <- coronavirus %>% filter(cases==max(cases)) 
goofy

ggplot(data = df3,
       mapping = aes(x = date,
                     y = cumulative_deaths,
                     color = country)) +
  geom_line()
ggplot(data = df3,
       mapping = aes(x = date,
                     y = rate,
                     color = country)) +
  geom_line()
summary(df3)
library(scales)
library(ggrepel)
library(glue)
library(lubridate)
as_of_date <- df3 %>% 
  summarise(max(date)) %>% 
  pull()
as_of_date_formatted <- glue("{wday(as_of_date, label = TRUE)}, {month(as_of_date, label = TRUE)} {day(as_of_date)}, {year(as_of_date)}")

ggplot(data = df3,
       mapping = aes(x = date, 
                     y = cumulative_cases, 
                     color = country)) +
  # represent cumulative cases with lines
  geom_line(size = 0.7, alpha = 0.8) +
  # add points to line endings
  geom_point() +
  # add country labels, nudged above the lines
  # geom_label_repel(nudge_y = 1, direction = "y", hjust = 1) + 
  # turn off legend
  guides(color = FALSE) +
  # use pretty colors
  scale_color_viridis_d() +
  # better formatting for y-axis
  scale_y_continuous(labels = label_comma()) +
  # use minimal theme
  theme_minimal() +
  # customize labels
  labs(
    x = "My pretty Lab 3 visual",
    y = "Cumulative number of deaths",
    title = "Cumulative deaths from COVID-19, selected countries",
    subtitle = glue("Data as of", as_of_date_formatted, .sep = " "),
    caption = "Source: github.com/RamiKrispin/coronavirus"
  )
```

